### 2013-02-13 doc/xps2013/proposal.tex
modified:   proposal.bib 	modified:   proposal.tex

\paragraph{Hardware support for aggregation:} If the increased latency caused
by extra hops in hierarchical aggregation turns out to be a big factor, we
will explore reducing by move some of the aggregation logic to the network
card, either by exploiting off-the-shelf programmable network cards or
experimental systems such ask NetFPGA~\cite{netfpga} (while we are planning to
focus on Infiniband-based networking, if we explore the NetFPGA route we will
be using Ethernet).

@@

Networking infrastructure supports remote read and write operations.  Read operations take a source address and key (on a remote node), a destination address and key (on the local node), and a length.  Cards support address translation for regions of memory that are registered with them.  These translations, indexed by the supplied key, are used to perform virtual to physical address translation.  The physical address of the destination address is similarly translated using the translation table at the local node.  Cards also support remote write operations with a similar set of semantics, where addresses are translated on cards using the cards local translation table.  Both read and write operations are asynchronous.  The requesting processor initiates the operation and then can check for completion by querying a completion queue at the local card.  \emph{Notably}, the processor(s) on the remote node are provided a polling interface whereby they can learn if a remote write operation occurred.  No such interface exists to learn about remote read or atomic operations.

@@

%% The second half of this paragraph is really confusing, I have just a minimal clue of what it is talking about. And the little I understand sounds a bit vague and not directly related to grappa.
The second challenge is the virtual to physical address translation on the cards requires the operating system to pin translations.  This removes paging flexibility from the operating system because runtime systems are (rationally) built to expect data structures to be in memory capable of RDMA or not.  Those in RDMA memory must be pinned.  In the long run, the solution here likely lays at the intersection of what progress is occurring in processor support for virtual machines and virtual memory translation for I/O.  Our goal will be to carefully study the I/O virtualization proposals put forth by major processor manufactures and devise solutions that enable RDMA to capitalize on these developments.

@@

Grappa applications are likely to use large amounts of data, 100s of terabytes to petabytes. Keeping all these data in DRAM only is unlikely to be cost effective. We will explore mechanisms that allow data to reside in solid-state storage {\em only}.

@@

\noindent\textbf{Year 1:} The first year of this effort will focus on creating the benchmark suite and initial language and compiler support.  We aim to pull together 6-8 benchmarks, sample data sets, and scripts to measure performance results on the initial Grappa runtime, UPC~\cite{upc} implementations, and serial reference code.  For the compiler we will have the front end done enabling easy addition of keywords to annotate data-types and structural elements (parallel for, etc).  The compiler will be generating in-efficient Grappa code by the end of the first year.

>>>>

\paragraph{Hardware support for aggregation:} If the increased latency caused
by extra hops in hierarchical aggregation turns out to be detrimental, we
will explore mitigation by moving some of the aggregation logic to the network
card, either by exploiting off-the-shelf programmable network cards or
experimental systems such ask NetFPGA~\cite{netfpga} (while we are planning to
focus on Infiniband-based networking, if we explore the NetFPGA route we will
be using Ethernet).

@@

Networking infrastructure supports remote read and write operations.  Read operations take a source address and key (on a remote node), a destination address and key (on the local node), and a length.  Cards support address translation for regions of memory that are registered with them.  These translations, indexed by the supplied key, are used to perform virtual to physical address translation.  The physical address of the destination address is similarly translated using the translation table at the local node.  Cards also support remote write operations with a similar set of semantics, where addresses are translated on a card using the card's local translation table.  Both read and write operations are asynchronous.  The requesting processor initiates the operation and then can check for completion by querying a completion queue at the local card.  \emph{Notably}, the processor(s) on the remote node are provided a polling interface whereby they can learn if a remote write operation occurred.  No such interface exists to learn about remote read or atomic operations.

@@

%% The second half of this paragraph is really confusing, I have just a minimal clue of what it is talking about. And the little I understand sounds a bit vague and not directly related to grappa.
The second challenge is that the virtual to physical address translation on the cards requires the operating system to pin translations.  This removes paging flexibility from the operating system, because runtime systems are (rationally) built to expect data structures to be in memory capable of RDMA or not.  Those in RDMA memory must be pinned.  In the long run, the solution here likely lies at the intersection of what progress is occurring in processor support for virtual machines and virtual memory translation for I/O.  Our goal will be to study carefully the I/O virtualization proposals put forth by major processor manufactures and devise solutions that enable RDMA to capitalize on these developments.

@@

Grappa applications are likely to use large amounts of data, 100s of terabytes to petabytes. Keeping all these data in DRAM only is unlikely to be cost effective. We will explore mechanisms that use solid-state storage as primary memory, greatly expanding the directly addressible storage at low cost.

@@

\noindent\textbf{Year 1:} The first year of this effort will focus on creating the benchmark suite and initial language and compiler support.  We aim to pull together 6-8 benchmarks, sample data sets, and scripts to measure performance results on the initial Grappa runtime, UPC~\cite{upc:2005} implementations, and serial reference code.  For the compiler we will have the front end done enabling easy addition of keywords to annotate data-types and structural elements (parallel for, etc).  The compiler will be generating in-efficient Grappa code by the end of the first year.
